---
title: 'Reproducible Research: Peer Assessment 1'
output:
  html_document:
    keep_md: yes
---


## Loading and preprocessing the data
Have set Working Directory to the Cloned Folder  in Global Options

```{r Loading and Transform the Data}
     unzip("./activity.zip")
     activityData<-read.csv("./activity.csv",header =TRUE,
                            stringsAsFactors =FALSE)
```


## What is mean total number of steps taken per day?
**Task1 : Histogram of Total Steps**  
Calcultaing the Number of Steps per Day and making a Histogram  
Using Library dplyr
```{r dailysteps}
     library(dplyr,quietly=TRUE)
     steps<-group_by(activityData,date)
     dailySteps<-summarise(steps,total_Steps =sum(steps))
     hist(dailySteps$total_Steps,main="Distribution of Daily Steps",
          xlab ="Cummulative Daily Steps",ylab ="Avg Steps")
```

Printing Mean and Median of the Daily Steps
```{r dailysteps Mean and Median}
     mean(dailySteps$total_Steps,na.rm =TRUE)
     median(dailySteps$total_Steps,na.rm =TRUE)
```

## What is the average daily activity pattern?
**Task 2: Average  Daily Activity Pattern and interval of Max Activity**

```{r Avg Activity Pattern}
interval_activity<-group_by(activityData,interval)
     avg_steps<-summarise(interval_activity,
                             int_avg_steps =mean(steps,na.rm =TRUE))
     with(avg_steps,plot(interval,int_avg_steps,type ="l",
                          main=" Avg Daily Activity Pattern",xlab ="Interval"))
```

Interval of Maximum Activity
```{r}
avg_steps[avg_steps$int_avg_steps==max(avg_steps$int_avg_steps),]$interval
```


## Imputing missing values

**Task 3: Input Missing Values using interval medians, Histogram**  
Number of Missing Values
```{r Input mising Values using Insert Function}
     count(activityData[is.na(activityData$steps),])
```

Creating a Median Steps per interval Data Frame
```{r}
   median_steps<-summarise(interval_activity,
    int_median_steps =median(steps,na.rm =TRUE))
```

Function for Inserting Data from the interval Median Summary
```{r Function for replacing NAs}
  insertData<-function(df11,df22)
    {for(index in 1:nrow(df11))
     {if(is.na(df11[index,"steps"]))
        {df11[index,"steps"]<-(df22[df22$"interval"== df11[index,"interval"],])$int_median_steps}
              
    } 
   df11
}
```
Filling in the missing Values into a new Data Frame and Making a Histogram
```{r Analysis using median of the intervals for NAs}
   activityDataComplete<-insertData(activityData,median_steps)
## count(activityDataComplete[is.na(activityDataComplete$steps),])
   stepsModified<-group_by(activityDataComplete,date)
   dailySteps1<-summarise(stepsModified,total_Steps =sum(steps))
   hist(dailySteps1$total_Steps,main="Distribution of Daily Steps(NAs inserted)", xlab ="Cummulative Daily Steps")
     a<- mean(dailySteps$total_Steps,na.rm =TRUE) 
     b<- mean(dailySteps1$total_Steps,na.rm =TRUE)
     c<- median(dailySteps$total_Steps,na.rm =TRUE) 
     d<- median(dailySteps1$total_Steps,na.rm =TRUE)
```
Table of  Mean and Median for Original and Complete Data  
It stays the nearly Same if I were to insert intevl means.
```{r}
   matrix(c(a,b,c,d),2,2, dimnames = list(c("Original Data Set","Completed Data Set"),c("Mean","Median")))
```

## Are there differences in activity patterns between weekdays and weekends?
*Task 4 Activity Patterns using an additonal Column and Panel Printing*
```{r Adding additional factor Column and then Comparing}
   activityData$wend <-as.factor(ifelse(weekdays(as.Date(activityData$date))   %in%   c("Saturday","Sunday"), "Weekend", "Weekday"))
      steps<-group_by(activityData,wend,interval)  
      weeklysteps<-summarise(steps,avg_steps=mean(steps,na.rm =TRUE))
      library(lattice)
      xyplot(weeklysteps$avg_steps ~ weeklysteps$interval | weeklysteps$wend, 
                layout = c(1, 2),type ="l",xlab ="intervals",ylab="steps")
```
##End of Assignment